---
title: 'Statistical tests'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## One sample Wilcoxon Test

Wilcoxon's rank sum test (also known as the unpaired Wilcoxon rank sum test or the Mann-Whitney U test)	It is the test for ordinal or continuous data. In contrast to Student's t-test, does not require the data to be normally distributed. This test too can be used for paired or unpaired data.

Dataset shows different method for identifying protiens and the accuracy.
```{r}
load("rdas/protStruct.rda")
head(protStruct)
```

```{r}
str(protStruct)
```

Test against a specific hypothesis, mean =0.5
```{r}
wilcox.test(protStruct$Correct,mu=0.5) 
```

The results of this test concur with the results of the t-test that the central measure (mean or median) of the data differs significantly from 0.5. 

## Two sample Wilcoxon Rank Sum Tests
Load the data

Dataset includes info on effectivesness of 2 painkillers as rated by patients
```{r}
load("rdas/medrank.rda")
medrank
```

```{r include=FALSE}
library(dplyr)
```

```{r}
group_by(medrank, Drug) %>%
  summarise(
    count = n(),
    medi_score = median(Score, na.rm = TRUE),
    IQR_score = IQR(Score, na.rm = TRUE),
    medi_rank = median(Rank, na.rm = TRUE),
    IQR_rank = IQR(Rank, na.rm = TRUE)
  )
```

```{r}
oxycodone_rank <- medrank$Rank[medrank$Drug=="Oxycodone"]
ibuprofen_rank <- medrank$Rank[medrank$Drug=="Ibuprofen"]
```

Test the 2 drugs
```{r}
relation <- wilcox.test(oxycodone_rank, ibuprofen_rank)
relation
```

No difference between the median of 2 groups


## Unpaired Wilcoxon Rank Sum Tests

Load the data
```{r}
load("rdas/my_data.rda")
head(my_data)
```

```{r}
group_by(my_data, group) %>%
  summarise(
    count = n(),
    median = median(weight, na.rm = TRUE),
    IQR = IQR(weight, na.rm = TRUE)
  )

```

```{r include=FALSE}
library("ggpubr")
```

```{r}
ggboxplot(my_data, x = "group", y = "weight", 
          color = "group", palette = c("#00AFBB", "#E7B800"),
          ylab = "Weight", xlab = "Groups")
```

###  Compute two-samples Wilcoxon test - Method 1

```{r}
women_weight <- my_data$weight[my_data$group=="Woman"]
men_weight <- my_data$weight[my_data$group=="Man"]
res <- wilcox.test(women_weight, men_weight)
res
```


### Compute two-samples Wilcoxon test - Method 2

```{r}
res <- wilcox.test(weight ~ group, data = my_data,
                   exact = FALSE)
res
```

To test whether the median men's weight is less than the median women's weight:
```{r}
wilcox.test(weight ~ group, data = my_data, 
            exact = FALSE, alternative = "less")
```

To test whether the median men's weight is greater than the median women's weight
```{r}
wilcox.test(weight ~ group, data = my_data,
            exact = FALSE, alternative = "greater")
```


## Pairwise Wilcoxon Rank Sum Tests

Load the data
```{r}
attach(airquality)
head(airquality)
```

```{r}
Month <- factor(Month, labels = month.abb[5:9])
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
pairwise.wilcox.test(Ozone, Month)
```
These give warnings because of ties

```{r message=FALSE, warning=FALSE}
pairwise.wilcox.test(Ozone, Month, p.adj = "bonf")
detach()
```


## Krusal Wallis Test

Load the data

Dataset is about the reaction time taken after intake of 3 different drinks
```{r}
load("rdas/reactionR.rda")
str(reactionR)
```

Medians
```{r}
tapply(reactionR$ReactionTime, reactionR$Drink, median)
```

Summary
```{r}
Water_summary<-summary(reactionR$ReactionTime[reactionR$Drink=='Water'])
Coffee_summary<-summary(reactionR$ReactionTime[reactionR$Drink=='Coffee'])
Alcohol_summary<-summary(reactionR$ReactionTime[reactionR$Drink=='Alcohol'])
```

```{r}
compare1<-cbind(Water_summary,Coffee_summary,Alcohol_summary)
round(compare1,2)
```

```{r}
boxplot(reactionR$ReactionTime~reactionR$Drink)
```

```{r}
kruskal.test(reactionR$ReactionTime~reactionR$Drink)
```

The simplest adjustment is the Bonferroni adjustment p.adj='bonferroni' which multiplies each Wilcoxon signed rank p-value by the total number of Wilcoxon tests being carried out (here it is 3) while the exact=F stands for the asymptotic test which allows tied ranks.

Look for which drinks are

```{r}
pairwise.wilcox.test(reactionR$ReactionTime,reactionR$Drink,p.adj='bonferroni',exact=F)
```

## Friedman Test

Load the data
```{r}
load("rdas/Data.rda")
str(Data)
```

```{r}
Data$Likert.f <- as.factor(Data$Likert)
XT <- xtabs( ~ Instructor + Likert.f,
             data = Data)
XT
```

As proportions
```{r}
prop.table(XT, 
           margin = 1)
```

```{r include=FALSE}
library(lattice)
```

```{r}
histogram(~ Likert.f | Instructor,
          data=Data,
          layout=c(1,5)      #  columns and rows of individual plots
)
```


```{r}
friedman.test(Likert ~ Instructor | Rater,
              data = Data)

```

Also done as
```{r}
friedman.test(Data$Likert, Data$Instructor, Data$Rater)
```

```{r}
Ratertable <- xtabs(Likert ~ Instructor + Rater, 
                    data = Data)
```

Can do rowmeans of ratertable to see the difference in mean ratings,the diff is tested statistically by friedmantest

### Conover test

```{r include=FALSE}
library(PMCMR)
```

```{r}
PT = posthoc.friedman.conover.test(y      = Data$Likert,
                                   groups = Data$Instructor,
                                   blocks = Data$Rater,
                                   p.adjust.method="fdr")
PT
```

