---
title: "Statistical Tests"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chi-square Test

Load the data
```{r}
load("rdas/research_funding_rates.rda") 
```
```{r}
research_funding_rates
```

Plot research funding success rates: Total, Men, Women
```{r}
boxplot(research_funding_rates$success_rates_total,research_funding_rates$success_rates_men, 
     research_funding_rates$success_rates_women, xlab=c("Total, Men, Women"), 
     main ="Research funding success rates")
```

```{r}
summary(research_funding_rates)
```

Compute the totals that were successful and the totals that were not as follows:
```{r include=FALSE}
library(dplyr)
```

```{r}
totals <- research_funding_rates %>%  select(-discipline) %>% summarize_all(funs(sum)) %>%  
  summarize(yes_men = awards_men, no_men = applications_men - awards_men, yes_women = awards_women,              
            no_women = applications_women - awards_women)
totals
```

Larger percent of men than women received awards
```{r}
totals %>% summarize(percent_men = yes_men/(yes_men+no_men),                      
                       percent_women = yes_women/(yes_women+no_women)) 
```

Percent funding when randomly assigned
```{r}
funding_rate <- totals %>%  summarize(percent_total = 
                    (yes_men + yes_women)/(yes_men + no_men +yes_women + no_women)) %>% .$percent_total
funding_rate
```

Chi-square test is to compare any signficant difference between the groups with the help of a two-by-two table

Create the two-by-two Observed data table:
```{r}
Observed_data <- data.frame(awarded = c("no", "yes"),                       
                           men = c(totals$no_men, totals$yes_men),                      
                           women = c(totals$no_women, totals$yes_women))
Observed_data
```

Have a look at the two-by-two expected data table:
```{r}
Expected_data <- data.frame(awarded = c("no", "yes"),      
             men = (totals$no_men + totals$yes_men) * c(1 - funding_rate, funding_rate),        
             women = (totals$no_women + totals$yes_women)*c(1 - funding_rate, funding_rate))
Expected_data
```

The idea is to compare the observed to expected.
```{r}
chisq_test <- Observed_data %>% select(-awarded) %>% chisq.test()
chisq_test
qchisq(0.95,df=1)
```

H0 is There is no difference beween the groups

H1 is there is difference between the groups

Pvalue is >0.05 at 95% significance level. So Ho cannot be rejected.

The Chisquare value of 3.8111 wont cross the critical value at 95% signifcance level for df=1 is 3.841459, so H0 cannot be rejected

Conclusion is that there is no difference between the groups

### Odds Ratio
```{r}
odds_men <- (Observed_data$men[2] / sum(Observed_data$men)) / (Observed_data$men[1] / sum(Observed_data$men))
odds_women <- (Observed_data$women[2] / sum(Observed_data$women)) / (Observed_data$women[1] / sum(Observed_data$women))
odds_ratio <- odds_men / odds_women
odds_ratio
```

Larger sample size will give lesser p-value but odds ratio remains same.

### Confidence interval for Odds Ratio

```{r}
log_or <- log( odds_men / odds_women ) 
se <- Observed_data %>%  select(-awarded) %>% summarize(se = sqrt(sum(1/men) + sum(1/women))) %>%  .$se 
ci <- log_or + c(-1,1) * qnorm(0.975) * se 
ci
```



## Fisher's Test
Similar to chisquare test used when sample size is very small <5

Fisher's test with lady tasing tea for milk poured before or after tea was poured.
```{r}
tab <- matrix(c(3,1,1,3),2,2) 
rownames(tab)<-c("Poured Before","Poured After") 
colnames(tab)<-c("Guessed before","Guessed after") 
tab
```

Idea is test actual to guess really works or was it just random happening
```{r}
fisher.test(tab)
fisher.test(tab, alternative="greater") 
```

H0 is that there is no difference

H1 is that there is difference between groups

p-value is >0.05 so H0 cannot be rejected. It cannot be said with statistical sig that the lady can actucally guess.

When sample size increase p-vlaue gets reduced but odd ratio remains same.



## McNemar Test

McNemar test for paired data
```{r}
x<-matrix(c(21,9,2,12),2,2)
mcnemar.test(x)
mcnemar.test(x,correct=FALSE)
```

The argument correct = FALSE is a logical indicating whether to apply continuity correction when computing the test statistic.

```{r, include=FALSE}
library(exact2x2)
```

```{r}
mcnemar.exact(x)
```
